---
title: "Representations of what’s possible reflect others’ epistemic states"

date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cosmo
    highlight: tango
---

```{r, message=FALSE, warning=FALSE, include = FALSE}
# Load packages
library("kableExtra") 
library("pander")
library("knitr")
library("tidyverse")
library("stringr")
library("effectsize")
library("lme4") 
library("emmeans")
library("afex")
library("janitor")
```

## Study 1: Possibility Generation

#Read in Data

```{r, importing data, message=FALSE, warning=FALSE, include = FALSE}
## Cleaning Data and Prepare Dataframes for Demographics and Analysis

######################################   Read in data from study 1
df.data1 <- read.csv("study1.csv") %>% 
  filter(row_number() > 2) %>% # additional rows in qualtrics csv 
  clean_names() %>% 
    filter(!status =="Survey Preview", #exclude bot fails,
         #!str_detect(start_date, "1/26/24"), # Exclude rows where start_date contains "1/26/24"
         progress ==100| progress==99)

######################################   Demographics

df.desc1<-df.data1 %>% 
  select(duration_in_seconds, age, ethnicity, ethnicity_6_text, gender, ethnicity, education)%>% 
    mutate(duration_in_seconds=as.numeric(duration_in_seconds)) %>% 
    mutate(age= as.numeric(age)) %>% 
    summarise(n=n(),
             mean_age=mean(age),
             sd_age=sd(age),
             female=sum(gender=="Female"),
             male=sum(gender=="Male"),
             other=sum(gender=="Other"),
             white=sum(ethnicity=="White / Caucasian"),
             black=sum(ethnicity=="Black / African American"),
             latino=sum(ethnicity=="Hispanic / Latino"),
             native=sum(ethnicity=="Native American / American Indian"),
             asian=sum(ethnicity=="Asian / Pacific Islander"),
             other_e=sum(ethnicity=="Other"),
             mean_time=mean(duration_in_seconds/60),
             sd_time=sd(duration_in_seconds/60)) 

#participant_counts <- df.exp1 %>%
#  group_by(scenario, epistemic_condition) %>%
#  summarise(number_of_participants = n_distinct(participant)) %>% 
#  ungroup() %>% 
#            summarise(mean = mean(number_of_participants))

#print(participant_counts)


  
```

# Edit Data

```{r, warning=FALSE, message=FALSE, include=FALSE}
# Contuinued Data Editing for Main Analysis 

df.exp1 = df.data1 %>% 
  mutate(participant = 1:n()) %>% 
  select(s1_fr_k_4:s16_a5_i_6,
         participant) %>%
    select(-contains("last_click"), -contains("click_count")) %>% 
  gather(condition, response, -participant) %>% 
  filter(response != "") %>% 
  mutate(scenario = case_when(
    str_detect(condition, "s1_") ~ "scenario 1",
    str_detect(condition, "s2_") ~ "scenario 2",
    str_detect(condition, "s3_") ~ "scenario 3",
    str_detect(condition, "s4_") ~ "scenario 4",
    str_detect(condition, "s5_") ~ "scenario 5",
    str_detect(condition, "s6_") ~ "scenario 6",
    str_detect(condition, "s7_") ~ "scenario 7",
    str_detect(condition, "s8_") ~ "scenario 8",
    str_detect(condition, "s9_") ~ "scenario 9",
    str_detect(condition, "s10_") ~ "scenario 10",
    str_detect(condition, "s11_") ~ "scenario 11",
    str_detect(condition, "s12_") ~ "scenario 12",
    str_detect(condition, "s13_") ~ "scenario 13",
    str_detect(condition, "s14_") ~ "scenario 14",
    str_detect(condition, "s15_") ~ "scenario 15",
    str_detect(condition, "s16_") ~ "scenario 16")) %>% 
mutate(epistemic_condition = case_when(
    str_detect(condition, "_k_") ~ "Knowledge",
    str_detect(condition, "_i_") ~ "Ignorance")) %>% 
  mutate(responsetype = case_when(
    str_detect(condition, "s4_a3_i_7")  ~ "normal_resp_3", #this one was weirdly coded
    str_detect(condition, "fr_k_4") |  str_detect(condition, "fr_i_4")  ~ "resp_1",
    str_detect(condition, "fr_k_5") |  str_detect(condition, "fr_i_5")  ~ "resp_2",
    str_detect(condition, "fr_k_6") |  str_detect(condition, "fr_i_6")  ~ "resp_3",
    str_detect(condition, "fr_k_7") |  str_detect(condition, "fr_i_7")  ~ "resp_4",
    str_detect(condition, "fr_k_8") |  str_detect(condition, "fr_i_8")  ~ "resp_5",
    str_detect(condition, "a1_k_4") |  str_detect(condition, "a1_i_4")  ~ "prob_resp_1",
    str_detect(condition, "a1_k_5") |  str_detect(condition, "a1_i_5")  ~ "moral_resp_1",
    str_detect(condition, "a1_k_6") |  str_detect(condition, "a1_i_6")  ~ "normal_resp_1",
    str_detect(condition, "a2_k_4") |  str_detect(condition, "a2_i_4")  ~ "prob_resp_2",
    str_detect(condition, "a2_k_5") |  str_detect(condition, "a2_i_5")  ~ "moral_resp_2",
    str_detect(condition, "a2_k_6") |  str_detect(condition, "a2_i_6")  ~ "normal_resp_2",
    str_detect(condition, "a3_k_4") |  str_detect(condition, "a3_i_4")  ~ "prob_resp_3",
    str_detect(condition, "a3_k_5") |  str_detect(condition, "a3_i_5")  ~ "moral_resp_3",
    str_detect(condition, "a3_k_6") |  str_detect(condition, "a3_i_6")  ~ "normal_resp_3",
    str_detect(condition, "a4_k_4") |  str_detect(condition, "a4_i_4")  ~ "prob_resp_4",
    str_detect(condition, "a4_k_5") |  str_detect(condition, "a4_i_5")  ~ "moral_resp_4",
    str_detect(condition, "a4_k_6") |  str_detect(condition, "a4_i_6")  ~ "normal_resp_4",
    str_detect(condition, "a5_k_4") |  str_detect(condition, "a5_i_4")  ~ "prob_resp_5",
    str_detect(condition, "a5_k_5") |  str_detect(condition, "a5_i_5")  ~ "moral_resp_5",
    str_detect(condition, "a5_k_6") |  str_detect(condition, "a5_i_6")  ~ "normal_resp_5",
    str_detect(condition, "a1_k_7") |  str_detect(condition, "a1_i_7")  ~ "prob_resp_1",
    str_detect(condition, "a1_k_8") |  str_detect(condition, "a1_i_8")  ~ "moral_resp_1",
    str_detect(condition, "a1_k_9") |  str_detect(condition, "a1_i_9")  ~ "normal_resp_1",
    str_detect(condition, "a2_k_7") |  str_detect(condition, "a2_i_7")  ~ "prob_resp_2",
    str_detect(condition, "a2_k_8") |  str_detect(condition, "a2_i_8")  ~ "moral_resp_2",
    str_detect(condition, "a2_k_9") |  str_detect(condition, "a2_i_9")  ~ "normal_resp_2",
    str_detect(condition, "a3_k_7") |  str_detect(condition, "a3_i_7")  ~ "prob_resp_3",
    str_detect(condition, "a3_k_8") |  str_detect(condition, "a3_i_8")  ~ "moral_resp_3",
    str_detect(condition, "a3_k_9") |  str_detect(condition, "a3_i_9")  ~ "normal_resp_3",
    str_detect(condition, "a4_k_7") |  str_detect(condition, "a4_i_7")   ~ "prob_resp_4",
    str_detect(condition, "a4_k_8") |  str_detect(condition, "a4_i_8")   ~ "moral_resp_4",
    str_detect(condition, "a4_k_9") |  str_detect(condition, "a4_i_9")  ~ "normal_resp_4",
    str_detect(condition, "a5_k_7") |  str_detect(condition, "a5_i_7")  ~ "prob_resp_5",
    str_detect(condition, "a5_k_8") |  str_detect(condition, "a5_i_8")  ~ "moral_resp_5",
    str_detect(condition, "a5_k_9") |  str_detect(condition, "a5_i_9")  ~ "normal_resp_5",
    str_detect(condition, "first_click") ~ "first_click",
    str_detect(condition, "page_submit") ~ "page_submit")) %>% 
    select(-condition) %>% 
    pivot_wider(names_from = responsetype,
              values_from = response) 
  
#is.list(df.exp1$probability) 

write.csv(df.exp1, "df.exp1.csv", row.names = FALSE)


```

```{r}

```



## Create Long Dataframe


```{r}
df.exp1_long = df.exp1 %>% 
    gather(responsetype, response, -scenario, -participant, -epistemic_condition, -first_click, -page_submit) %>% 
 mutate(response_order = case_when(
    str_detect(responsetype, "_1") ~ "1",
    str_detect(responsetype, "_2") ~ "2",
    str_detect(responsetype, "_3") ~ "3",
    str_detect(responsetype, "_4") ~ "4",
    str_detect(responsetype, "_5") ~ "5")) %>% 
  mutate(response_category = case_when(
    str_detect(responsetype, "prob_") ~ "probability",
    str_detect(responsetype, "moral_") ~ "morality",
    str_detect(responsetype, "normal_") ~ "normality",
    TRUE ~ "generated response")) %>% 
    select(-responsetype) 


```

## Create wide dataframe

```{r}
df.exp1_wide = df.exp1_long %>% 
  pivot_wider(names_from = response_category,
              values_from = response) %>% 
  mutate(across(c(probability, morality, normality), as.numeric)) %>% 
    mutate(avg=(normality+probability+morality)/3)

```

## Create long dataframe with average ratings across normality, probablity, morality

```{r}
df.exp1_long_avg = df.exp1_wide %>% 
    mutate(across(c(probability, morality, normality, avg), ~as.numeric(as.character(.)))) %>%
  pivot_longer(
    cols = c(probability, morality, normality, avg), 
    names_to = "response_category", 
    values_to = "response"
  )

```


#Export dataset with generated actions and ratings for further analysis

```{r}

df.exp1_long_exp <- df.exp1_long_avg  %>% 
  filter(response_category != "generated response")

write.csv(df.exp1_long_exp, "df.exp1_actions.csv", row.names = FALSE)

```

#Plot rating distributions in Knowledge and Ignorance condition (just to check)

```{r}
figS0 <- df.exp1_long_avg%>% 
 mutate(question = fct_recode(response_category,
                               value = "avg"), # Recoding 'avg' to 'value'
    question = factor(question, levels = c("normality", "morality", "probability", "value"))) %>% 
  filter(scenario =="scenario 7") %>% 
    ggplot(aes(x = response, fill = question, color = question)) +
      geom_histogram(position = "identity",
                     alpha = 0.33,
                     bins = 15) +
   facet_grid(question ~ epistemic_condition)

figS0

```
# Calculate Proportion of Best Option 
`
```{r}
# percentage of time that a participant rates each answer for a given scenario highest. ties go to the first answer


df.perBest <- df.exp1_wide %>% 
  select("participant", "scenario", "epistemic_condition", "response_order", "avg") %>%
  group_by(participant, scenario) %>% 
  mutate(rank = rank(-avg, ties.method = "first")) %>% 
  group_by(response_order, epistemic_condition) %>% 
  summarise(numBest = sum(rank == 1), .groups = "drop") %>% 
  group_by(epistemic_condition) %>% # Group by epistemic_condition
  mutate(perBest = numBest / sum(numBest)) %>% 
  ungroup()

```

#Create Dataframe with just knowledge condition

```{r}
df.exp1_wide_k <-  df.exp1_wide %>% 
  filter(epistemic_condition=="Knowledge")
```


# Here we predict the order in which an answer was generated based on how highly the answer is rated in the knowledge condition

```{r}
#  anova predicting answer number by average rating of an answer
pgPredictAnova <- summary(aov(lm(as.numeric(response_order) ~ avg, data=df.exp1_wide_k)))


lmerSamp_Full <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) +  scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_k)
    
    # lesion out probability
    lmerSamp_P <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_k)
    # lesion out morality
    lmerSamp_M <- lmer(as.numeric(response_order) ~ scale(probability) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_k)
    # lesion out normality
      lmerSamp_N <- lmer(as.numeric(response_order) ~ scale(morality) + scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_k)
      
  pgPredictP.df <- anova(lmerSamp_Full,lmerSamp_P)
  pgPredictM.df <- anova(lmerSamp_Full,lmerSamp_M)
  pgPredictN.df <- anova(lmerSamp_Full,lmerSamp_N)
```

#Create Dataframe with just ignorance condition

```{r}
df.exp1_wide_i <-  df.exp1_wide %>% 
  filter(epistemic_condition=="Ignorance")
```


# Here we predict the order in which an answer was generated based on how highly the answer is rated in the ignorance condition

```{r}

pgPredictAnova <- summary(aov(lm(as.numeric(response_order) ~ avg, data=df.exp1_wide_i)))


lmerSamp_Full <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) +  scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_i)
    
    # lesion out probability
    lmerSamp_P <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_i)
    # lesion out morality
    lmerSamp_M <- lmer(as.numeric(response_order) ~ scale(probability) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_i)
    # lesion out normality
      lmerSamp_N <- lmer(as.numeric(response_order) ~ scale(morality) + scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide_i)
      
  pgPredictP.df <- anova(lmerSamp_Full,lmerSamp_P)
  pgPredictM.df <- anova(lmerSamp_Full,lmerSamp_M)
  pgPredictN.df <- anova(lmerSamp_Full,lmerSamp_N)
```


# Both conditions together: Here we predict the order in which an answer was generated based on how highly the answer is rated

```{r}
#  anova predicting answer number by average rating of an answer
pgPredictAnova <- summary(aov(lm(as.numeric(response_order) ~ avg, data=df.exp1_wide)))


lmerSamp_Full <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) +  scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide)
    
    # lesion out probability
    lmerSamp_P <- lmer(as.numeric(response_order) ~ scale(morality) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide)
    # lesion out morality
    lmerSamp_M <- lmer(as.numeric(response_order) ~ scale(probability) + scale(normality) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide)
    # lesion out normality
      lmerSamp_N <- lmer(as.numeric(response_order) ~ scale(morality) + scale(probability) + 
                         (1|scenario) + 
                         (scale(morality) + scale(normality) +  scale(probability) |participant),
                       data=df.exp1_wide)
      
  pgPredictP.df <- anova(lmerSamp_Full,lmerSamp_P)
  pgPredictM.df <- anova(lmerSamp_Full,lmerSamp_M)
  pgPredictN.df <- anova(lmerSamp_Full,lmerSamp_N)
  

  
   # predicting normality from interaction of morality and probability
  lmerNorm_Full <- lmer(scale(normality) ~ scale(morality) * scale(probability) + (1|scenario) +
                   (scale(probability) * scale(morality) |participant), data=df.exp1_wide)
  # predicting normality with morality and probability as separate fixed effects
  lmerNorm_Main <- lmer(scale(normality) ~ scale(morality) + scale(probability) + (1|scenario) +
                   (scale(probability) * scale(morality) |participant), data=df.exp1_wide)
  # model with probability removed
  lmerNorm_P <- lmer(scale(normality) ~ scale(morality) + (1|scenario) +
                   (scale(probability) * scale(morality) |participant), data=df.exp1_wide)
  # model with morality removed
  lmerNorm_M <- lmer(scale(normality) ~ scale(probability) + (1|scenario) +
                   (scale(probability) * scale(morality) |participant), data=df.exp1_wide)
  

  pgNormInt <- anova(lmerNorm_Full,lmerNorm_Main) # effect of interaction
  
  pgNormP <- anova(lmerNorm_Main,lmerNorm_P) # effect of probability
  
  pgNormM <- anova(lmerNorm_Main,lmerNorm_M) # effect of morality
   
  
```

# Figure showing Ratings of 5 generated actions by epistemic conditions (in CogSci paper)

```{r fig1, echo=F, warning=F, message=F, fig.width=6.5,fig.height=5.25}
#install.packages("viridis")  # Install
library("viridis")           # Load

fig1 <- df.exp1_long %>% 
  mutate(epistemic_condition = factor(epistemic_condition, levels = c("Knowledge", "Ignorance"))) %>%
  filter(response_category != "generated response") %>% 
  mutate(response_category = factor(response_category)) %>%
  mutate(response = as.numeric(response)) %>%
  mutate(response_order = factor(response_order)) %>%
  ggplot(aes(x = response_order, y = response, fill = response_category)) +
  coord_cartesian(xlim = c(0.75, 5.25), ylim = c(0, 100)) +
  geom_boxplot(position = "dodge", alpha = .4, outlier.alpha = 0, width=0.8, notch = TRUE) +
  geom_point(aes(color = response_category), position = position_jitterdodge(jitter.width = .35), alpha = 0.05) +
  labs(
    x = "Action Generation number", 
    y = "Value of option generated", 
    fill = "Rating:", 
    color = "Rating:"
  ) +
  geom_smooth(
    aes(y = response, x = as.numeric(response_order), color = response_category), 
    method = "lm", 
    position = position_dodge(width = .75)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    legend.position = "top",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = rel(1.4)),
    axis.title = element_text(size = rel(1.5), vjust = .9),
    axis.text.x = element_text(size = rel(1.5)),
    axis.ticks = element_blank(),
    legend.text = element_text(size = rel(1.3)),
    legend.title = element_text(size = rel(1.3)),
    strip.text = element_text(size = rel(1.3)),  # Matches font size in facet panels
    strip.background = element_rect(fill = "white", color = "black")  # New panel background color
  ) +
scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) +
  facet_wrap(~ epistemic_condition)

ggsave(fig1, file = "figMNPratings.pdf", width = 11, height = 7)

fig1
```
# Figure Proportion of Best Option by Action Generation Number (in Cogsci paper)

```{r}
figS1 <- df.perBest %>%
  mutate(epistemic_condition = factor(epistemic_condition, levels = c("Knowledge", "Ignorance"))) %>%
  ggplot(aes(x = as.numeric(response_order), y = perBest, color = epistemic_condition, group = epistemic_condition)) + 
  geom_point(size =2.5) +
  geom_line(size = 1.3) +
  labs(x = "Generation number", y = "Proportion of options generated that were best", color = "Condition") +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    legend.position = "top",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = rel(1.5)),
    axis.title = element_text(size = rel(1.5), vjust = .9),
    axis.text.x = element_text(size = rel(1.5)),
    axis.ticks = element_blank(),
    legend.text = element_text(size = rel(1.5)),
    legend.title = element_text(size = rel(1.5)),
    strip.text = element_text(size = rel(1.5)),
    strip.background = element_rect(fill = "white", color = "black")
  )+
    scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) 

print(figS1)
ggsave(figS1, file = "figBest.pdf", width = 7, height = 7)
```


